Which of my Transient Type Checks are not (Almost) Free?

### 2019-09-11 paper.tex
Run-time (at the time of running). Runtime (the engine supporting execution)
Signed-off-by: Stefan Marr <git@stefan-marr.de>

Although na\"ive implementations of transient type checks have a high runtime overhead, just-in-time compilation and optimisation in virtual machines can eliminate much of this overhead.  Unfortunately
the improvement is not uniform: while most type checks can be optimised
away, some will significantly decrease a program's
performance, and some may even increase it.

@@

This has also lead to a technique to evaluate the performance of gradual typing, such as the so called so-called ``Takikawa'' protocol \cite{Takikawa2016,Greenman2019jfp,vitousek-transient-arXive-2019}. The Takikawa protocol was created to measure the runtime cost of gradual typing by testing various configurations of typed/untyped code. This approach was designed to investigate the the performance effect of the amount of typed vs untyped code, in particular it shows that how simply adding more type annotations does not always produce a uniform effect on performance. Here we adapt their approach in order to identify individual type annotations that may be responsible for performance effects.

@@

The high-performance computing community has been investigating how
tools and visualisations can help developers to utilise their systems
more efficiently \citep{Papenhausen:2016:IVT,daSilva:2019:PSV}.
Their focus is typically on parallelisation opportunities,
guided by runtime feedback, cost models, or heuristics. 
Their large body of work \citep{Isaacs:2014:PerfViz} uses various approaches,
though, we are not aware of work that has used an approach similar to ours.

At the moment, our approach to identifing type annotations that case
performance anomalies is not integrated into a development environment.
Though, for instance Optimization Coaching
\citep{St-Amour:2012:OCO} is a promising direction.
Optimization Coaching uses feedback from the runtime to guide developers
to insert or change type declarations 
to enable a compiler to generate a more optimal program.
In this spirit, we would eventually want to achieve the same, although
in our case,
we need to run full experiments to get the necessary information.

@@

In the future, we hope to investigate statistical techniques to
determine the significance of each type annotation's contribution to a
programs overall performance. We would also like to investigate 
whether this approach can assist with optimsations for programmers' day-to-day  development, or help VM engineers identifying performance bugs in the underlying virtual machines.

>>>>

Although na\"ive implementations of transient type checks have a high run-time overhead, just-in-time compilation and optimisation in virtual machines can eliminate much of this overhead.  Unfortunately
the improvement is not uniform: while most type checks can be optimised
away, some will significantly decrease a program's
performance, and some may even increase it.

@@

This has also lead to a technique to evaluate the performance of gradual typing, such as the so called so-called ``Takikawa'' protocol \cite{Takikawa2016,Greenman2019jfp,vitousek-transient-arXive-2019}. The Takikawa protocol was created to measure the run-time cost of gradual typing by testing various configurations of typed/untyped code. This approach was designed to investigate the the performance effect of the amount of typed vs untyped code, in particular it shows that how simply adding more type annotations does not always produce a uniform effect on performance. Here we adapt their approach in order to identify individual type annotations that may be responsible for performance effects.

@@

The high-performance computing community has been investigating how
tools and visualisations can help developers to utilise their systems
more efficiently \citep{Papenhausen:2016:IVT,daSilva:2019:PSV}.
Their focus is typically on parallelisation opportunities,
guided by run-time feedback, cost models, or heuristics. 
Their large body of work \citep{Isaacs:2014:PerfViz} uses various approaches,
though, we are not aware of work that has used an approach similar to ours.

At the moment, our approach to identifying type annotations that case
performance anomalies is not integrated into a development environment.
Though, for instance Optimization Coaching
\citep{St-Amour:2012:OCO} is a promising direction.
Optimization Coaching uses feedback from the runtime to guide developers
to insert or change type declarations 
to enable a compiler to generate a more optimal program.
In this spirit, we would eventually want to achieve the same, although
in our case,
we need to run full experiments to get the necessary information.

@@

In the future, we hope to investigate statistical techniques to
determine the significance of each type annotation's contribution to a
programs overall performance. We would also like to investigate 
whether this approach can assist with optimisations for programmers' day-to-day  development, or help VM engineers identifying performance bugs in the underlying virtual machines.
